---
title: "Week 8"
author: "Igor Serafini"
date: "2023-03-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(github_link)[]

#Assignment 8 Metabarcoding & Community Ecology

Questions: 

1. What effect (if any) does garlic mustard have on the plant community?

2.What has a stronger effect on plant communities: the presence/absence of garlic mustard (in/out) or sampling population? 


Loading packages

```{r}
library(ggplot2)
library(ape)
library(ggtree)
library(vegan)
library(tidyverse)
library(reshape2)
```

Loading data
```{r}
Floristic <- read.csv("FloristicSurvey.csv")
```

Cleaning data
```{r}
#Retaining useful data, and using Population and Location for pivot longer 
Floristic1 <- Floristic %>% 
  select(-Quadrate,-Bud_Flw, -Flower, -Flw_Sil, -Sillique, -Bolting, -Rosettes) %>% 
  pivot_longer(cols= c(-Location, Population),names_to = "name", values_to = "value") %>% 
  group_by(Location, name) %>% 
  mutate(row_id = row_number()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = "name", values_from = "value", values_fn = list) %>% 
  unnest_wider(everything(), names_sep = "_")
```


Further arranging data
```{r}
Floristic1_selected1 <- Floristic1 %>% 
  select(4:38)

Floristic_selected2 <- Floristic1 %>% 
  select(Location_1, Population_1)

result1 <- cbind(Floristic1_selected1, Floristic_selected2)

result <- result1 %>% 
pivot_longer(cols = c(-Location_1, -Population_1))
```



Species into 30 rows for NMDS 
```{r}
Species<- data.frame(names(Floristic1))

Species <- head(Species, 30)

#Note, unlisting Species data
Species <- data.frame(matrix(unlist(Species)))
```


Setting Florsitic data as matrix
```{r}
floristic1 <- result1 %>% 
  select(-Location_1, -Population_1)


floristic <- as.matrix(floristic1)
```


Running NMDS, setting distance to bray 
```{r}
nmds <- metaMDS(floristic1, distance= "bray")

plot(nmds)
```

```{r}
data.scores0 <- as.data.frame(scores(nmds)$sites)
```


Adding Specie collumn into data.score0 (NMDS site data) back
```{r}
data.scores0$Species <- Species$matrix.unlist.Species..
head(data.scores0)
```

Running NMDS 
```{r}
NMDS_plot <- ggplot(data.scores0, aes(x = NMDS1, y = NMDS2)) + 
    geom_point(size = 4, aes(colour= Species)) + geom_text(aes(label = Species), size = 3, hjust = 0.2, vjust = -1) 
NMDS_plot
```
Note, the cluster of Budding (garlic mustard is visible at near the coordinate NMDS2= ~-1.3 and NMDS1= ~0. In this context, we can see that garlic mustard has a close composition to Claytonia virginia, however, a very low composition to the other species. Thus, we can run other models that will better help us understand the impact of garlic mustard. 


Running a Permonova to establish the 
```{r}
species_matrix <- as.matrix(result1[,1:35])

# indicating presence of in or out = Location- Code just ran better not sure why 
result1$GarlicMustard <- ifelse(result1$Budding > 0, "in", "out")

#PERMANOVA
permanova1 <- adonis2(species_matrix ~ GarlicMustard + Population_1, data = result1, method = "bray")

# results
permanova1
```
ANS 1. The permanova demonstrates that Garlic mustard has a significant effect on community composition, explianing 14.8% of its variation
ANS 2. The permonova demonstrates that Sampliing population has a larger effect on plant communities, such that Garlic mustard presence/absence account for 14.8% of variability in the population, while Sampling Population Accounts for 45.2%





